Practical Optional Types for Clojure

### 2016-01-14 intro.tex
colors, footnote to examples

\begin{figure*}[t!]
  \normalsize
\begin{lstlisting}
(*typed ann pname [(U File String) -> (U nil String)] typed*)
(defmulti pname class)  ; multimethod dispatching on class of argument
(defmethod pname String [s] (*invoke pname (*interop new File s interop*) invoke*)) ; String case 
(defmethod pname File [f] (*interop .getName f interop*)) ; File case, static null check
(*invoke pname "STAINS/JELLY" invoke*) ;=> "JELLY" :- (U nil Str)
\end{lstlisting}
\caption{A simple Typed Clojure program (delimiter colors: {\color{interop}Java interoperation (green)}, 
  {\color{types}type annotation (blue)},
  {\color{invoke}function invocation (black)}, {\color{red}collection literal (red)}, {\color{mygray}miscellaneous (gray)})}
\label{fig:ex1}
\end{figure*}

>>>>

\begin{figure*}[t!]
  \normalsize
\begin{lstlisting}
(*typed ann pname [(U File String) -> (U nil String)] typed*)
(defmulti pname class)  ; multimethod dispatching on class of argument
(defmethod pname String [s] (*invoke pname (*interop new File s interop*) invoke*)) ; String case 
(defmethod pname File [f] (*interop .getName f interop*)) ; File case, static null check
(*invoke pname "STAINS/JELLY" invoke*) ;=> "JELLY" :- (U nil Str)
\end{lstlisting}
\caption{A simple Typed Clojure program (colors: {\color{interop}Java interoperation (green)}, 
  {\color{types}type annotation (blue)},
  {\color{invoke}function invocation (black)}, {\color{red}collection literal (red)}, {\color{mygray}other (gray)})}
\label{fig:ex1}
\end{figure*}

### 2016-01-14 overview.tex
change colors, fix breaks
colors, footnote to examples

We now begin a tour of the central features of Typed Clojure,
beginning with Clojure itself. Our presentation
uses the full Typed Clojure system to illustrate key type system
ideas, before studying the core features in detail in
\secref{sec:formal}.

@@

@@

The \clj{:Meal} entry is common to each HMap, always mapped to a known keyword singleton
type.
It's natural to dispatch on the \clj{class} of an instance---it's similarly
natural to dispatch on a known entry like \clj{:Meal}.
%
\begin{exmp2} % use other environment to work around page break
\begin{lstlisting}
(*typed ann desserts [Order -> Int] typed*)
(defmulti desserts :Meal)  ; dispatch on :Meal entry
(defmethod desserts :lunch [o] (*invoke :desserts d invoke*))
(defmethod desserts :dinner [o] (*invoke :desserts d invoke*))
(defmethod desserts :combo [o] 
  (*invoke + (*invoke desserts (*invoke :meal1 o invoke*) invoke*) (*invoke desserts (*invoke :meal2 o invoke*) invoke*) invoke*))

(*invoke desserts {*map :Meal :combo, :meal1 {*map :Meal :lunch :desserts 1 map*}, 
           :meal2 {*map :Meal :dinner :desserts 2 map*} map*} invoke*) ;=> 3
\end{lstlisting}
\label{example:desserts-on-meal}
\end{exmp2}
%
The \clj{:combo} method is verified to only structurally recur
on \clj{Order}s. This is achieved because we learn the argument \clj{o} must
be of type
\clj{'\{:Meal :combo\}}
since
\clj{(isa? (:Meal o) :combo)}
is true. Combining this
with the fact that \clj{o} is an \clj{Order}
eliminates possibility of \clj{:lunch} and \clj{:dinner}
orders, simplifying \clj{o} to
\clj{'\{:Meal ':combo :meal1 Order :meal2 Order\}}
which contains appropriate arguments for both recursive calls.

>>>>

We now begin a tour of the central features of Typed Clojure,
beginning with Clojure itself. Our presentation
uses the full Typed Clojure system to illustrate key type system
ideas,\footnote{Full examples: \url{https://github.com/typedclojure/esop16}} before studying the core features in detail in
\secref{sec:formal}.

@@

\newpage

@@

The \clj{:Meal} entry is common to each HMap, always mapped to a known keyword singleton
type.
It's natural to dispatch on the \clj{class} of an instance---it's similarly
natural to dispatch on a known entry like \clj{:Meal}.

\newpage

\begin{exmp}
\begin{lstlisting}
(*typed ann desserts [Order -> Int] typed*)
(defmulti desserts :Meal)  ; dispatch on :Meal entry
(defmethod desserts :lunch [o] (*invoke :desserts d invoke*))
(defmethod desserts :dinner [o] (*invoke :desserts d invoke*))
(defmethod desserts :combo [o] 
  (*invoke + (*invoke desserts (*invoke :meal1 o invoke*) invoke*) (*invoke desserts (*invoke :meal2 o invoke*) invoke*) invoke*))

(*invoke desserts {*map :Meal :combo, :meal1 {*map :Meal :lunch :desserts 1 map*}, 
           :meal2 {*map :Meal :dinner :desserts 2 map*} map*} invoke*) ;=> 3
\end{lstlisting}
\label{example:desserts-on-meal}
\end{exmp}
%
The \clj{:combo} method is verified to only structurally recur
on \clj{Order}s. This is achieved because we learn the argument \clj{o} must
be of type
\clj{'\{:Meal :combo\}}
since
\clj{(isa? (:Meal o) :combo)}
is true. Combining this
with the fact that \clj{o} is an \clj{Order}
eliminates possibility of \clj{:lunch} and \clj{:dinner}
orders, simplifying \clj{o} to
\clj{'\{:Meal ':combo :meal1 Order :meal2 Order\}}
which contains appropriate arguments for both recursive calls.
